Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-03 content/impl.tex
started to move proofs to appendix; removing dependency on minted

We will next show, in Lemma~\ref{lem:sound}, that our analysis is sound, i.e. that the result of an IDE problem obtained through a correlated-calls transformation removes only facts that occur on infeasible paths. To prove the Soundness Lemma, we first introduce Lemmas~\ref{lem:sound1} and~\ref{lem:sound3}.

We will denote the top element in the environment lattice as
\begin{equation}
  \Omega=\lambda d\,.\,\topcc\,.
\end{equation}

For the purpose of the proofs, we will rewrite Equation~\eqref{eq:edgefndef} that defines an edge function as follows:
  \begin{equation}\label{eq:edgefnThroughDelta}
    \ccedgefn S=\lambda e\,.\,
    \begin{cases}
      \id&\text{if $d_1=d_2=\Lambda$,}\\
      \lambda m\,.\,\varepsilon(e)(\delta(m))&\text{otherwise},
    \end{cases}
  \end{equation}
  where $S\subseteq R$, $d_1$ and $d_2$ are the source and target facts, and for a map $m\in\lcc_U$, $\delta(m)$ is either $m$ or $\botcc$:
  \begin{equation}\label{eq:deltadef}
    \delta(m)=\begin{cases}
      \botcc&\text{if $d_1=\Lambda$}\\
      m&\text{otherwise.}
    \end{cases}
  \end{equation}

Additionally, for a path $p=[\startmain,\,\dots]$ and a fact $d\in D$, we will denote the lattice element that is mapped to $d$ according to the flow functions of path $p$ as follows:
\begin{equation}
  \mpd=\menv(p)(\Omega)(d)\,.
\end{equation}

The following Lemma shows that the lattice elements (receiver-to-types maps) of a correlated-calls IDE analysis correctly overapproximate the possible types of a receiver in a program execution.

\begin{lemma}\label{lem:sound1}
Let $p=[\startmain,\,\dots,\,n]$ be some concrete execution trace of the program, and let $r\in R$ be a receiver. If after the execution trace $p$, at node $n$, $r$ points to an object of runtime type~$t$, and $d\in D$ is a fact such that $d\in M_F(p)(\varnothing)$, then
  \begin{equation}
    t\in\mpd(r)\,.
  \end{equation}
\end{lemma}
\begin{proof}
  By induction on the length of the trace.

  \textit{Basis:} $p=[\startmain]$. Then there is no instruction at which a receiver $r$ could be instantiated, and the Lemma is trivially true.

  \textit{Induction hypothesis:} Let $p=[\startmain,\,\dots,\,n_{k-1}]$, and let $\uptau$ be the set of types to which $\mpdkm$ maps $r$:
  \begin{equation}
    \uptau=\mpdkm(r)\,.
  \end{equation}
  Assume that for a concrete execution path $p=[\startmain,\,\dots,\,n_{k-1}]$, at node $(n_{k-1},\,d_{k-1})$, the Lemma holds, i.e. $t\in\uptau$.

  \textit{Induction step:} Let $p'=[\startmain,\,\dots,n_{k-1},\,n_k]$ and $t'\in T$ be the type to which $r$ is mapped at $n_k$.

  For each $i$, let $e_i$ be the edge $((n_{i-1},\,d_{i-1}),\,(n_i,\,d_i))$. Note that $$e_1=((\startmain,\,\Lambda),\,(n_1,\,d_1))\,.$$

  Observe that
  \begin{align*}
    \mppd
    &=\menv(p')(\Omega)(d)\\
    &=\left(\menv(e_k)\circ\menv(e_{k-1})\circ\ldots\circ\menv(e_1)\right)(\Omega)(d)\\
    &=\menv(e_k)\left(\menv(p)(\Omega)\right)(d)\,.
  \end{align*}

  According to~\eqref{eq:envTransToEdgeFnEdge},
  \begin{align*}
    &\menv(e_k)\left(\menv(p)(\Omega)\right)(d)(r)\\
      =&\bigg(\ccedgefn R((n_{k-1},\,\Lambda),\,(n_k,\,d))(\topcc)\sqcap\\
       &\bigsqcap_{d'\in D}\ccedgefn R((n_{k-1},\,d'),\,(n_k,\,d))(\menv(p)(\Omega)(d'))\bigg)(r)\\
      \supseteq&
        \bigsqcap_{d'\in D}\ccedgefn R((n_{k-1},\,d'),\,(n_k,\,d))(\menv(p)(\Omega)(d'))(r)\\
      \supseteq&\,\ccedgefn R((n_{k-1},\,d_{k-1}),\,(n_k,\,d))(\mpdkm)(r)\,.
  \end{align*}
  Therefore, 
  \begin{equation}\label{eq:musubset}
    \efek\subseteq\mppd(r)\,.
  \end{equation}

  We will now show that
  \[
    t'\in\efek\,,
  \]
  which, due to~\eqref{eq:musubset}, means that the Lemma holds.

  According to~\eqref{eq:edgefnThroughDelta}, there are two cases in which $\ccedgefn R(e_k)$ could fall.

  If $d_{k-1}=d_k=\Lambda$, then $d_k\notin M_F(p)(\varnothing)$, since it does not belong to the set $D$, and the Lemma trivially holds.

  Otherwise, 
  \[
    \ccedgefn R(e_k)=\lambda m\,.\varepsilon(e_k)(\delta(m))\,.
  \]
  It follows that
  \begin{align}\label{eq:efek}
    \efek
    &=(\lambda m\,.\varepsilon(e_k)(\delta(m)))(\mpdkm)(r)\notag\\
    &=\varepsilon(e_k)(\delta(\mpdkm))(r).
  \end{align}
  Let us denote the lattice element $\delta(\mpdkm)$ with $\Delta$:
  \[
    \Delta=\delta(\mpdkm)\,.
  \]
  Note that since $\Delta$, according to~\eqref{eq:deltadef}, can be either $\botcc$ or $\mpdkm$, it always maps $r$ to a set containing~$t$:
  \begin{equation}\label{eq:deltaContainsT}
    t\in\Delta(r)\,.
  \end{equation}
  Note also that unless the instruction at $n_{k-1}$ contains an assignment for $r$, $r$ is mapped to the same object of type $t$ as at node $n_{k-1}$, and $t=t'$. Therefore, for the non-assignment instructions, it is sufficient to prove that $t\in\Delta(r)$.

   Depending on the instructions at the nodes $n_{k-1}$ and $n_k$, there are four cases:
  \begin{enumerate}
    \item\label{item:asgn} The instruction at $n_{k-1}$ is an assignment for a receiver $r'\in R$.
        Since $\varepsilon_R(e_k)=\lambda m\,.\,m[r'\to\bot_T]$,
        \begin{align*}
          \efek&=            
            (\lambda m\,.\,m[r'\to\bot_T])(\Delta)(r)\\
          &=\Delta[r'\to\bot_T](r)\,.
        \end{align*}
      In the resulting map, $r'$ is mapped to $\bot_T$. Then

      \begin{enumerate}
        \item if $r=r'$, then $\efek=\bot_T$, which contains $t'$.
        \item\label{item:defaultmap} If $r\ne r'$, then $r$ has not been reassigned a value, and still maps to the same object of type $t$. The receiver $r$ is mapped to $\Delta(r)$, which, according to~\eqref{eq:deltaContainsT}, contains $t$. Since $t=t'$, $\Delta(r)$ contains $t'$.
      \end{enumerate}
    \item\label{item:callstart} $e_k$ is a call-start edge with signature $s_\mathcal F$, and $f\in\mathcal F$ is the called procedure.
      Then
        \begin{align*}
          \efek
          &=(\lambda m\,.\,m[r'\to m(r')\cap\tau(s_\mathcal F,\,f)])(\Delta)(r)\\
          &=\Delta[r'\to\Delta(r')\cap\tau(s_\mathcal F,\,f)]\,,
        \end{align*}
      where $r'$ is the receiver of the call.
      \begin{itemize}
        \item If $r'=r$, then $\Delta(r')=\Delta(r)$ which contains $t$. Since $t\in\tau(s_\mathcal F,\,f)$, it follows that $t\in\Delta(r)\cap\tau(s_\mathcal F,\,f)$, and $t\in\efek$.
         \item If $r'\ne r$, see~(\ref{item:defaultmap}).
       \end{itemize}
    \item $e_k$ is an end-return edge, $r_1,\,\dots,\,r_k\in R$ are the local variables in the callee method, $r'$ is the receiver of the call site corresponding to the return node $n_k$, and $f\in\mathcal F$ is the called method with signature~$s_\mathcal F$.
      Then 
      \[
        \varepsilon_R(e_k)=\lambda m\,.\,m[r'\to m(r')\cap\tau(s_\mathcal F,\,f)][r_1\to\bot_T]\dots[r_k\to\bot_T].
      \]
      If $r\in\{r_1,\,\dots,\,r_k\}$, see~Case~\ref{item:asgn}. Otherwise, the case is analogous to Case~\ref{item:callstart}.
    \item\label{item:idcase} The node contains any other instruction.
      Then 
      \[
        \ccedgefn R(e_k)(\mpdkm)(r)=\id(\Delta)(r)=\Delta(r),
      \]
      which contains $t$ according to~\eqref{eq:deltaContainsT}.\qedhere
  \end{enumerate}
\end{proof}

We will now show that on a node of a concrete execution path, the correlated-calls analysis does not map receivers to $\top_T$. In other words, the analysis never considers nodes of a concrete execution path unreachable.

\begin{lemma}\label{lem:sound3}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, $r\in R$ a receiver, and $d\in D$ a data-flow fact. Then if $d\in M_F(p)(\varnothing)$,
  \begin{equation}
    \mpd(r)\ne\top_T\,.
  \end{equation}
\end{lemma}
\begin{proof}
  By induction on the length of the execution trace.

  \textit{Basis:} 
    Let $p=[\startmain]$. Since the only realizable path corresponding to $p$ is $[(\startmain,\,\Lambda)]$, there is no fact $d\in D$ such that $d\in M_F(p)(\varnothing)$, and the claim follows immediately.

  \textit{Induction hypothesis:} 
  Let $p=[\startmain,\,\dots,\,n_{k-1}]$. Let $\uptau$ be the set of types to which $r$ is mapped by $\mpdkm$:
  \begin{equation}
    \uptau=\mpdkm(r)\,.
  \end{equation}
  Assume the Lemma holds for that for a concrete execution path $$p=[\startmain,\,n_1,\,\dots,\,n_{k-1}]\,,$$ i.e. $\uptau\ne\top_T$ for an arbitrary $r\in R$ and $d_{k-1}\in D$.

  \textit{Induction step:}
    Let $p'=[\startmain,\,n_1,\,\dots,n_{k-1},\,n_k]$ be a concrete execution path.

    Let $e_k=((n_{k-1},\,d_{k-1}),\,(n_k,\,d))$. As shown in~\eqref{eq:musubset},
    \begin{align*}
      \mppd(r)
        &\supseteq\ccedgefn R(e_k)(\mpdkm)(r)\,.
    \end{align*}

    From Definition~\ref{def:edgefn}, we can see that unless $e_k$ is a call-start edge or an end-return edge, the result follows from the induction hypothesis.
    More formally, if $e_k$ is not a call-start or end-return edge, then for all $m\in\lcc_R$,
    \[
      \ccedgefn R(e_k)(m)\sqsubseteq m\,.
    \]
     The edge function corresponding to the call-start and end-return edges is the only place in which the set of types that a receiver maps to can be reduced.

    Assume that $e_k$ is a end-return edge with a call on the receiver $r'\in R$ with a signature $s_\mathcal F$ to a function $f\in\mathcal F$.
    \begin{align*}
      \ccedgefn R&(e_k)(\mpdkm)(r)\\
      &=\left(\lambda m\,.\,m[r'\to m(r)\cap\tau(s_\mathcal F,\,f)][r_1\to\bot_T]\dots[r_l\to\bot_T]\right)(\mpdkm)(r)\\
      &=\left(\mpdkm[r'\to \uptau\cap\tau(s_\mathcal F,\,f)][r_1\to\bot_T]\dots[r_l\to\bot_T]\right)(r)\,,
    \end{align*}
  where $r_1,\,\dots,r_l\in R$ are the local variables in the called method.

    If $r\in\{r_1,\,\dots,\,r_l\}$, then $\efek=\bot_T\ni t$\footnote{In the case of a recursive call, it is possible that both $r\in\{r_1,\,\dots,\,r_l\}$ and $r=r'$.
    In that case, the set to which $r$ will be mapped would be still ``overwritten'' by $\bot_T$.}.

    Otherwise, if $r=r'$, then $\efek=\uptau\cap\tau(s_\mathcal F,\,f)$.

    According to Lemma~\ref{lem:sound1} and by the induction hypothesis, the runtime type $t$ of $r$ must be contained in $\mpdkm(r)=\uptau$. At the same time, by definition, $t$ is part of $\tau(s_\mathcal F,\,f)$. Therefore, $t\in\uptau\cap\tau(s_\mathcal F,\,f)\subseteq\efek$, which means that $\efek\ne\top_T$.

    The same reasoning applies to the case where $e_k$ is a call-start edge.
\end{proof}

Finally, we will prove the soundness of the correlated-calls analysis: we will show that our analysis only considers a path infeasible if it cannot occur in a concrete execution of a program.

\begin{lemma}[Soundness]\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  \begin{equation}
    d\in \backCC\left(\result(\transCC_R(P))\right)(n)\,.
  \end{equation}
\end{lemma}
\begin{proof} Let $\rho=\result(\transCC_R(P))$. Then
  \begin{align*}
    \backCC(\rho)(n)
    &=D_n^\Subset(\rho)\\
    &=\left\{d'\ |\ 
        d'\in\mvp F(n)\,\wedge\,\forall r\in R:\ \rho(n,\ d')(r)\ne\top_T\right\}.
  \end{align*}
  Since $\mvp F(n)=\bigsqcap_{q\in\ivp(n)}M_F(q)(\varnothing)$, and $p\in\ivp(n)$, it follows that
    \begin{align*}
      d&\in M_F(p)(\varnothing)\\&\subseteq\mvp F(n)\,.
    \end{align*}
  At the same time, for all receivers $r\in R$,
  \begin{align*}
    \rho(n,\,d)(r)
    &=\left(\bigsqcap_{q\in\ivp(n)}\mpddef q d\right)(r)\\
    &=\bigsqcap_{q\in\ivp(n)}\mpddef q d(r)\,.
  \end{align*}
  According to~Lemma~\ref{lem:sound3}, $\mpd(r)\ne\top_T$.
  Since $p\in\ivp(n)$,
  \begin{align*}
    \mpd(r)\subseteq\bigsqcap_{q\in\ivp(n)}\mpddef qd(r)\,.
  \end{align*}
  From $\bigsqcap_{q\in\ivp(n)}\mpddef qd(r)=\rho(n,\,d)(r)$ it follows that
  $\mpd(r)\subseteq\rho(n,\,d)(r)$.
  Therefore, $\rho(n,\,d)(r)\ne\top_T$, and $d\in D_n^\Subset(\rho)=\backCC(\rho)(n)$.
\end{proof}

@@

We will describe a ``reduced'' correlated-calls transformation in which we only consider receivers from $\rcc$ and ignore other receivers of $R$. We will show that IDE problems obtained through ordinary and reduced correlated-calls transformations yield the same results.

The following Lemma shows that the types to which a given receiver is mapped in the result of the algorithm is not affected by other receivers and the types to which they are mapped.

\begin{lemma}\label{lem:recindepedgefn}
  Let $P$ be an IFDS problem. Let $N^*$ be the supergraph for $P$, $D$ the set of data-flow facts, $n\in N^*$ a node, and $p=[\startmain,\,\dots,\,n]$ a path in the supergraph. Let $d\in D\cup\{\Lambda\}$.
  Then for any realizable path $p'\in\textsf{RP}(p,\,d)$, set $S\subseteq R$, and receiver $r\in S$,
  \begin{equation}
    \ccedgefn S(p')(\topcc)(r)=
    \ccedgefn{\{r\}}(p')(\topcc)(r)\,.
  \end{equation}
\end{lemma}
\begin{proof}
  By induction on the length of $p$.

  \textit{Basis:} $p'=[(\startmain,\,\Lambda)]$. Then $\ccedgefn S(p')=\id=\ccedgefn{\{r\}}(p')$, and the Lemma follows directly.

  \textit{Induction hypothesis:} Suppose that for a path $q=[(\startmain,\,\Lambda),\,\dots,\,(n_{k-1},\,d_{k-1})]$, where $q\in\textsf{RP}(n,\,d)$, the Lemma holds, i.e. both edge functions map $r$ to the same set of types $\uptau$:
  \begin{align*}
    \uptau
    &=\ccedgefn S(q)(\topcc)(r)\\
    &=\ccedgefn{\{r\}}(q)(\topcc)(r)\,.
  \end{align*}

  \textit{Induction step:} Let $q'=[(\startmain,\,\Lambda),\,\dots,\,(n_{k-1},\,d_{k-1}),\,(n_k,\,d_k)]$ and the edge $e_k=((n_{k-1},\,d_{k-1}),\,(n_k,\,d_k))$.

  Observe that for any set $U\subseteq R$ such that $r\in U$,
  \begin{align}\label{eq:edgefnU}
    \ccedgefn U(q')(\topcc)(r)
    &=\ccedgefn U(e_k)(\ccedgefn U(q)(\topcc))(r)\,.
  \end{align}

  We can see from~\eqref{eq:edgefnThroughDelta} that there are two cases.

  If $d_{k-1}=d_k=\Lambda$, $\ccedgefn S(e_k)=\id=\ccedgefn{\{r\}}(e_k)$, and, due to~\eqref{eq:edgefnU},
  \begin{align*}
    \ccedgefn S(q')(\topcc)(r)&=\uptau\\
    &=\ccedgefn{\{r\}}(q')(\topcc)(r)\,.
  \end{align*}

  Otherwise, there are four sub-cases.
  \begin{enumerate}
    \item $e_k$ is a call-start edge, $r'.c()$ is the call site at $n_{k-1}$ with signature $s_\mathcal F$, $f\in\mathcal F$ is the called procedure, and $r'\in U$.
    Then
    \[
      \ccedgefn U(e_k)=\lambda m\,.\,\delta(m)[r'\to\delta(m)(r)\cap\tau(s_\mathcal F,\,f)]\,.
    \]
    There are two sub-cases.
    \begin{enumerate}
      \item\label{item:callstartreceq} If $r=r'$, then, according to~\eqref{eq:edgefnU}, the resulting set of types 
        \[
          \ccedgefn U(q')(\topcc)(r)=\delta(\ccedgefn U(q)(\topcc))(r)\cap\tau(s_\mathcal F,\,f).
        \]
        If $d_{k-1}=\Lambda$, then $\delta(\ccedgefn U(q)(\topcc))(r)=\botcc(r)=\bot_T$. If $d_{k-1}\ne\Lambda$, then $\delta(\ccedgefn U(q)(\topcc))(r)=\ccedgefn U(q)(\topcc)(r)=\uptau$. The set $\tau(s_\mathcal F,\,f)$ is the same for either case.

        Therefore, the value of $\ccedgefn U(q')(\topcc)(r)$ has the same result regardless of $U$,
        which means that $\ccedgefn S(q')(\topcc)(r)=\ccedgefn{\{r\}}(q')(\topcc)(r)$, and the Lemma holds.
      \item\label{item:callstartrecneq} If $r\ne r'$, then
        \begin{equation}
          \ccedgefn U(q')(\topcc)(r)=\delta(\ccedgefn U(q)(\topcc))(r)\,,
        \end{equation}
        which, as we have seen in~Case~\eqref{item:callstartreceq}, does not depend on~$U$, and the Lemma holds.
    \end{enumerate}
    \item $e_k$ is an end-return edge, $r_1,\,\dots,\,r_l\in U$ are the local variables in the callee method, $r'.c()$ is the call corresponding to the return node at $n_k$, $f\in\mathcal F$ is the called method with signature $s_\mathcal F$, and $r'\in U$.
    Then
    \[
      \ccedgefn U(e_k)=\lambda m\,.\,\delta(m)
      [r'\to\delta(m)(r)\cap\tau(s_\mathcal F,\,f)]
      [r_1\to\bot_T]\ldots[r_l\to\bot_T]\,.
    \]
    There are three sub-cases.
    \begin{enumerate}
      \item\label{item:localvarrec} If $r\in\{r_1,\,\dots,\,r_l\}$, then regardless of the value of $U$,
      \[
        \ccedgefn U(q')(\topcc)(r)=\bot_T\,,
      \]
      and the Lemma holds.
      \item Otherwise, if $r=r'$, the case is analogous to Case~\eqref{item:callstartreceq}.
      \item If $r\notin\{r',\,r_1,\,\dots,\,r_l\}$, then see Case~\eqref{item:callstartrecneq}.
    \end{enumerate}
    \item $n_{k-1}$ contains an assignment for $r'\in U$. Then
    \[
      \ccedgefn U(e_k)=\lambda m\,.\,\delta(m)[r'\to\bot_T]\,.
    \]
    If $r=r'$, see Case~\eqref{item:localvarrec}. If $r\ne r'$, see Case~\eqref{item:callstartrecneq}.
    \item Otherwise,
    \[
      \ccedgefn U(e_k)=\lambda m\,.\,\delta(m)\,,
    \]
    and the case is analogous to Case~\eqref{item:callstartrecneq}.\qedhere
  \end{enumerate}
\end{proof}

The following Lemma shows that the correlated-calls analysis computes the results for each receiver independently, or separately. To compute the set of types to which a receiver~$r$ is mapped at each exploded-graph node, we can exclude all other receivers in the program from the analysis (recall from~\eqref{eq:edgefndef} that the set of receivers that are considered in the analysis is specified by the set $S$ in a correlated-calls transformation $\transCC_S$). Therefore, for a given receiver $r$, the results of a $\transCC_S$- and a $\transCC_{\{r\}}$-analysis are the same.

\begin{lemma}\label{lem:recindep} Let $P$ be an IFDS problem. Let $N^*$ be the supergraph for $P$, $D$ the set of data-flow facts, and $S\subseteq R$ a set of receivers.
  Then for any $n\in N^*$, $d\in D$, and receiver $r\in S$,
  \begin{equation}
    \result\left(\transCC_S(P)\right)(n,\,d)(r)=
    \result(\transCC_{\{r\}}(P))(n,\,d)(r)\,.
  \end{equation}
\end{lemma}
\begin{proof}
  According to~\eqref{eq:ideresult}, \eqref{eq:mvpdef}, and~\eqref{eq:envTransToEdgefn},
  \begin{align}
    \result\left(\transCC_S(P)\right)(n,\,d)(r)
    &=\mvp{\textsf{Env}}(n,\,d)(r)\notag\\
    &=\left(\bigsqcap_{q\in\ivp(n)}M_\textsf{Env}(q)(\Omega)(d)\right)(r)\notag\\
    &=\left(\bigsqcap_{q\in\ivp(n)}\bigsqcap_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn S(q')(\topcc)\right)(r)\notag\\
    &=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn S(q')(\topcc)(r)\,.\label{eq:resultThroughEdgefn}
  \end{align}
  Then from Lemma~\ref{lem:recindepedgefn},
  \begin{align*}
    \result\left(\transCC_S(P)\right)(n,\,d)(r)
    &=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn{\{r\}}(q')(\topcc)(r)\\
    &=\result\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)\,.\qedhere
  \end{align*}
\end{proof}

The next lemma shows that the set of types to which a receiver is mapped in a correlated-calls lattice element can be represented as an intersection of static-type function applications $\tau(s_{\mathcal F_i},\,f_i)$.
\begin{lemma}\label{lem:edgefnThroughTaus}
  For an IFDS problem $P$, a node $n\in N^*$, and fact $d\in D$, let $p\in\mathsf{RP}(n,\,d)$ be a realizable path and $r\in R$ a receiver. Then there exists a non-negative number $\gamma$ of calls on the receiver $r$ with signatures $s_{\mathcal F_\gamma}$ to the functions $f_\gamma\in\mathcal F_\gamma$, for which
  \[
    \ccedgefn{\{r\}}(p)(\topcc)(r)=
      \bigcap_{\gamma\ge0}\tau(s_{\mathcal F_\gamma},\,f_\gamma)\,.
  \]
\end{lemma}
\begin{proof}
  Let $p$ have the following form\footnote{It can be shown from the definition of a pointwise representation in~Sagiv et al.~\cite{sagiv1996precise} that in a realizable path, there is never an edge from a fact of the set $D$ to a $\Lambda$ fact. Therefore, we can represent $p$ as a sequence of nodes that has a prefix of $\Lambda$-fact nodes, after which all nodes are non-$\Lambda$ facts.}:
  \[
    p=[(\startmain,\,\Lambda),\,(n_1,\,\Lambda),\,\dots,\,(n_k,\,\Lambda),
       (n_{k+1},\,d_{k+1}),\,\dots,\,(n_{k+l},\,d_{k+l})]\,,
  \]
  where $l\ge1$ and the facts for all nodes up to $n_k$ are equal to $\Lambda$ and $d_{k+i}\in D$ for $0<i\le l$.

  As previously, for all $i$, we will denote the edge $(n_i,\,n_{i+1})$ by $e_i$.

  From~\eqref{eq:edgefndef} we can infer that
  \[
    \ccedgefn{\{r\}}(p)=
    \ccedgefn{\{r\}}(e_{k+l})
    \circ\ldots
    \circ\ccedgefn{\{r\}}(e_{k+2})
    \circ(\lambda m\,.\,\beta)
    \circ\id
    \circ\ldots
    \circ\id\,,
  \]
  where
  \[
    \beta=
    \begin{cases}
      \botcc[r\to\tau(s_\mathcal F,\,f)]&\text{if $(n_k,\,n_{k+1})$ is a call-start or end-return edge, and}\\&\text{the call site $r.c()$ with signature $s_\mathcal F$ to the function}\\
      &\text{$f\in\mathcal F$ corresponds to the call-start or end-return edge,}\\
      \botcc&\text{otherwise\footnotemark.}
    \end{cases}
  \]
  \footnotetext{Since $d_k=\Lambda$ and $d_{k+1}\ne\Lambda$, the micro function for the edge $e_{k+1} $ is equal to $\lambda m\,.\,\varepsilon_{\{r\}}(e_{k+1})(\botcc)$. From the definition of $\varepsilon_S$~\eqref{eq:varepsilon} we can see that the only case where $\varepsilon_{\{r\}}(e_{k+1})(m)$ would not be equal to $\botcc$ is when $e_{k+1}$ is call-start or end-return edge.}

  Therefore,
  \begin{align}\label{eq:edgefnbeta}
    \ccedgefn{\{r\}}(p)(\topcc)
      &=\left(\ccedgefn{\{r\}}(e_{k+l})\circ\ldots\circ
        \ccedgefn{\{r\}}(e_{k+2})\right)((\lambda m\,.\,\beta)(\topcc))\notag\\
      &=\left(\ccedgefn{\{r\}}(e_{k+l})\circ\ldots\circ
        \ccedgefn{\{r\}}(e_{k+2})\circ\id\right)(\beta)\,.
  \end{align}

We can now prove the lemma by induction on $l$.

\textit{Basis:}
If $l=1$, then $\ccedgefn{\{r\}}(p)(\topcc)=\id(\beta)=\beta$.
There are two cases.

If $\beta=\botcc$, then 
\begin{align*}
  \ccedgefn{\{r\}}(p)(\topcc)(r)&=\beta(r)\\&=\bot_T\,,
\end{align*} and $\gamma=0$.

If $\beta=\botcc[r\to\tau(s_\mathcal F,\,f)]$, then 
\[
  \ccedgefn{\{r\}}(p)(\topcc)(r)=\tau(s_\mathcal F,\,f)\,,
\]
and $\gamma=1$.

\textit{Induction hypothesis:}
Assume that for a path $p=[(\startmain,\,\Lambda),\,\dots,\,(n_{k+l},\,d_{k+l})]$, the Lemma holds for $\gamma=N$, where $N\ge0$.

\textit{Induction step:}
Let $p'=[(\startmain,\,\Lambda),\,\dots,\,(n_{k+l},\,d_{k+l}),\,(n_{k+l+1},\,d_{k+l+1})]$.

Recall that
\begin{align*}
  \ccedgefn{\{r\}}(p')(\topcc)(r)
  &=\ccedgefn{\{r\}}(e_{k+l+1})\left(\ccedgefn{\{r\}}(p)(\topcc)\right)(r)\,.
\end{align*}

From~\eqref{eq:varepsilon} we can see that unless $e_{k+l+1}$ is a call-start or end-return edge corresponding to a call on the receiver $r$, then $\ccedgefn{\{r\}}(e_{k+l+1})(r)$ must be equal to either $\bot_T$ or $m(r)$, where $m=\ccedgefn{\{r\}}(p)(\topcc)$.

If $\ccedgefn{\{r\}}(e_{k+l+1})(r)=\bot_T$, then the Lemma holds for $\gamma=0$.

Otherwise,
\begin{align*}
  \ccedgefn{\{r\}}(e_{k+l+1})(\topcc)(r)&=\ccedgefn{\{r\}}(p)(\topcc)(r)\\
  &=\bigcap_{N}\tau(s_{\mathcal F_N},\,f_N)\,,
\end{align*}
and therefore $\gamma=N$.

Suppose that $e_{k+l+1}$ is a call-start edge with a call on the receiver $r$ with signature $s_\mathcal G$ to a function $g\in\mathcal G$. Then, according to~\eqref{eq:varepsilon}, 
\[
  \ccedgefn{\{r\}}(e_{k+l+1})=\lambda m\,.\,m[r\to m(r)\cap\tau(s_\mathcal G,\,g)]\,.
\]
Therefore,
\begin{align*}
  \ccedgefn{\{r\}}&(p')(\topcc)(r)\\
  &=\lambda m\,.\,m[r\to m(r)\cap\tau(s_\mathcal G,\,g)]\left(\ccedgefn{\{r\}}(p)(\topcc)\right)(r)\\
  &=\ccedgefn{\{r\}}(p)(\topcc)(r)\cap\tau(s_\mathcal G,\,g)\\
  &=\left(\bigcap_{N}\tau(s_{\mathcal F_N},\,f_N)\right)\cap\tau(s_\mathcal G,\,g)\,,
\end{align*}
and the Lemma holds for $\gamma=N+1$.

The case where $e_{k+l+1}$ is an end-return edge is analogous to the previous case.
\end{proof}

We now show that a receiver will be only mapped to $\topcc$ if it is the receiver of a correlated call.

\begin{lemma}\label{lem:ccrectop}
   For an IFDS problem $P$, let $n\in N^*$ be a node, and $d\in D$ a dataflow fact such that there exists a realizable path $p\in\textsf{RP}(n,\,d)$. Let $T$ be the set of all types in the program.
   If there exists a receiver $r\in R$ such that
  \[
    \ccedgefn{\{r\}}(p)(\topcc)(r)=\top_T\,,                       
  \]
  then $r\in \rcc$.
\end{lemma}
\begin{proof}
 According to Lemma~\ref{lem:edgefnThroughTaus},
 \[
   \ccedgefn{\{r\}}(p)(\topcc)(r)=
      \bigcap_{\gamma\ge0}\tau(s_{\mathcal F_\gamma},\,f_\gamma).
 \]

 Let $\tau_i=\tau(s_{\mathcal F_i},\,f_i)$. 
 For a given $k$, let $r.m_k()$ be the call site corresponding to $\tau_k$, and $T'$ the set of types compatible with the static type of $r$.
 Recall from Section~\ref{sec:ef} that 
 \begin{itemize}
   \item $\tau_k\ne\top_T$;
   \item if $\tau_k=T'$ then the corresponding call site is monomorphic;
   \item if $\tau_k\subset T'$ then the call site is polymorphic.
 \end{itemize}

 From the conditions of the Lemma, 
 \begin{equation}
   \bigcap_{\gamma\ge 0}\tau_\gamma=\top_T\,.
 \end{equation}

 If all $\tau_k=T'$, then $\bigcap_{\gamma\ge 0}\tau_\gamma$ is also equal to $T'$. Since $T'\ne\top_T$, this is a contradiction.

 If exactly one $\tau_k\subset T'$ and the rest are equal to $T'$, then $\bigcap_{\gamma\ge 0}\tau_\gamma$ is equal to $\tau_k$, which cannot be $\top_T$ either.

 Therefore, there are at least two sets, $\tau_i$ and $\tau_j$, which are strict subsets of $T'$. Since both $\tau_i$ and $\tau_j$ are non-empty and their intersection equals $\top_T$, $\tau_i$ and $\tau_j$ must be disjoint. If $\tau_i$ and $\tau_j$ are disjoint, they must correspond to different call sites.

 In other words, there are at least two calls on the same receiver for which the static-type function is a strict subset of the set of types compatible with a given receiver $r$. It follows that both calls have to be polymorphic. Therefore,~${r\in\rcc}$.
\end{proof}

We will now show that if a receiver ever gets mapped to top, then it is a correlated-calls receiver.

\begin{lemma}\label{lem:ccrectop}
   For an IFDS problem $P$, let $n\in N^*$ be a node, and $d\in D$ a dataflow fact such there exists a realizable path $p\in\textsf{RP}(n,\,d)$.
   Then, if there exists a receiver $r\in R$, such that
  \[
    \result\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)=\top_T\,,
  \]
  then $r\in \rcc$.
\end{lemma}
\begin{proof}
  As shown in~\eqref{eq:resultThroughEdgefn},
  \[
    \result\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn{\{r\}}(q')(\topcc)(r)\,.
  \]
  Since the latter is equal to $\top_T$, it follows that for each realizable path $p'$ to node $n$, $\ccedgefn{\{r\}}(p')(\top)(r)=\top_T$. According to~Lemma~\ref{lem:ccrectop}, this is only possible if $r\in\rcc$.
\end{proof}

Finally, we show that if a correlated calls analysis considers only correlated-call receivers, no precision is lost. A correlated-calls analysis that considers all receivers computes the same result as an analysis that considers only correlated-call receivers.

>>>>

We will next show that our analysis is sound, i.e. that the result of an IDE problem obtained through a correlated-calls transformation removes only facts that occur on infeasible paths.

\begin{lemma}[Soundness]\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  \begin{equation}
    d\in \backCC\left(\result(\transCC_R(P))\right)(n)\,.
  \end{equation}
\end{lemma}

@@

We will describe a ``reduced'' correlated-calls transformation in which we only consider receivers from $\rcc$ and ignore other receivers of $R$. We will show that IDE problems obtained through ordinary and reduced correlated-calls transformations yield the same results.
In other words, we show that if a correlated calls analysis considers only correlated-call receivers, no precision is lost.

\begin{lemma}\label{lem:onlycorrrec}
  Let $P$ be an IFDS problem. Then
  \begin{equation}
    \backCC\left(\result\left(\transCC_\rcc(P)\right)\right)=\backCC(\result\left(\transCC_R(P)\right))\,.
  \end{equation}
\end{lemma}
